\ttitle

### 2015-08-16 Chapters/CPET_Jaundice.tex
CPET-Jaundice - Major changes
- Reformat all tables
- Change all values to median (IQR)
- Change captions of all tables to fit with Donny style
- Include new binary log regression for vo2peak similar to vo2at
- Change all < and vo2 to correct tex format
- Next commit should be changes within chapter to reflect the changes in tables and more

\lhead{Chapter \ref{ch_cpet_jaundice}. \emph{CPET and Obstructive Jaundice}} % This is for the header on each page - perhaps a shortened title

@@

\section{Patients and Methods}
Patients who underwent classical or pylorus-preserving pancreaticoduodenectomy for periampullary lesions (both benign and malignant) between August 2008 and April 2013 and had undergone cardiopulmonary exercise testing as part of their preoperative workup at the West of Scotland Pancreatic Unit, Glasgow Royal Infirmary, Glasgow were included in the study. Established criteria for resectability in patients with malignant disease were used as outlined in previous published work. Segmental or wedge resection of the portal vein or superior mesenteric vein was carried out if the lesion was otherwise resectable.

@@

\subsection{Obstructive Jaundice}
Serum bilirubin levels were measured in all patients on the day before surgery. Obstructive jaundice (OJ) was defined as bilirubin levels greater than 35 micromol/litre and severe obstructive jaundice (sOJ) was defined as bilirubin levels greater than 250 micromol/litre. This threshold was selected because the DROP trial did not investigate patients with bilirubin levels greater than 250 micromol/litre and this study aimed to evaluate preoperative pathophysiology in this particular group.

Data on PBD (PBD) was also recorded. Serum bilirubin levels before and after biliary stenting were recorded [I will expand this section when I get the updated stent data]

\subsection{Cardiopulmonary Exercise Test}
Cardiopulmonary exercise tests were performed in the Department of Respiratory Medicine at the Glasgow Royal Infirmary using the ZAN-600 CPET suite (nSpire Health, Longmont, CO 80501, USA) (9). All patients underwent standard pulmonary function tests and spirometry prior to cardiopulmonary exercise testing. A cycle ergometer was used to perform a symptom-limited, incremental work-load test preceded by a 3-minute rest period. The test was stopped when patients achieved their maximum exercise tolerance, when significant ischaemic changes occurred on ECG or for other safety reasons. Peak oxygen consumption achieved at this stage was defined as VO$_2$Peak. The VO$_2$AT was calculated using the V-slope\parencite{beaver_new_1986,sue_metabolic_1988} and ventilatory equivalents\parencite{society_ats/accp_2003} methods. VO$_2$AT less than 10 ml/kg/min was considered to be low based on previous work by us\parencite{chandrabalan_pre-operative_2013} as well as Ausania and co-workers\parencite{ausania_effects_2012} which has shown increased incidence of complications in these patients. Oxygen consumption at peak exercise (VO$_2$Peak) was dichotomised using a cut-off of 16 ml/kg/min. Detailed description of cardiopulmonary exercise testing as well as the physiological parameters described in this study are published elsewhere.\parencite{balady_clinicians_2010}

\subsection{Statistics}
Grouping of the variables was carried out using standard or previously published thresholds. In the absence of such thresholds, the variables were treated as continuous variables. Non-parametric tests were used to analyse the association between categorical and continuous variables while Chi-square tests were used to analyse the association between categorical variables. Univariate and multivariate binary logistic regression analysis was used to study the relationship between preoperative patient characteristics and VO$_2$AT / VO$_2$Peak. SPSS software (Version 17.0; SPSS Inc., Chicago, IL, USA) was used to perform statistical analysis.

@@

\section{Results}
One-hundred and thirty eight patients had undergone pancreaticoduodenectomy (n=138), with preoperative cardiopulmonary exercise testing during the study period. Over half the patients were male (n=93, 67\%). Approximately half the number of patients were over the age of 65 (n=68, 49\%) and overweight or obese (n=69, 50\%). Cardiovascular comorbidity was present in 58 patients (42\%) and respiratory comorbidity was present in 12 patients (9\%). Fifty patients (36\%) had a history of cigarette smoking. The POSSUM Physiology Score was greater than 14 in 61 patients (44\%). Obstructive jaundice (serum bilirubin 35 – 250) was present in 32 (23\%) patients while severe obstructive jaundice (serum bilirubin > 250) was present in 19 (14\%) patients. 
The baseline demographic and clinical characteristics of non-jaundiced and jaundiced patients are shown in Table \ref{table:cpet_oj_table1}. A larger proportion of jaundiced patients were females compared to the non-jaundiced cohort (p<0.05) and smokers (p<0.05). Patients with jaundice were more likely to have an elevated POSSUM Physiology Score (p<0.005). Patients with cancer were more likely to be jaundiced (p<0.001). However, there was no statistically significant difference in age, BMI, cardiovascular comorbidity, or respiratory comorbidity between the non-jaundiced and jaundiced patients.

\input{Tables/c3_t1}

The relationship between obstructive jaundice and preoperative blood tests is shown in Table \ref{table:cpet_oj_table2}. While obstructive jaundice was statistically associated with multiple haematological and biochemical abnormalities, most of these did not appear to be of clinically significance. As expected, obstructive jaundice was associated with markedly elevated liver enzymes with severity of derangement associated with severity of jaundice. Obstructive jaundice and sOJ were associated with increasing CRP levels (p<0.001) and decreasing serum albumin levels (p<0.001). Obstructive jaundice was not associated with deranged renal function with both urea and creatinine remaining similar across all cohorts (p=0.09 and p=0.22 respectively).

\input{Tables/c3_t2}

There was no association between obstructive jaundice and preoperative pulmonary function tests (Table \ref{table:cpet_oj_table3}).

\input{Tables/c3_t3}

\subsubsection{Univariate analysis of obstructive jaundice versus CPET}
The relationship between obstructive jaundice and multiple physiological parameters measured at cardiopulmonary exercise testing is shown in Table \ref{table:cpet_oj_table4}. There was an inverse relationship between oxygen consumption at the anaerobic threshold (VO$_2$AT) and increasing severity of jaundice (p<0.05). However, no such linear relationship was noted between any of the other parameters measured both at anaerobic threshold and at peak exercise in spite of apparent statistically significant associations.

\input{Tables/c3_t4}

\subsubsection{Association between preoperative clinico-pathological factors and VO$_2$AT}
On multivariate analysis female sex (HR 3.75 CI 1.57-8.95 p<0.005), high BMI (HR 3.65 CI 1.61-8.26 p<0.005), presence of cancer (HR 4.02 CI 1.33-12.16 p<0.05) and raised CRP (HR 2.98 CI 1.29-6.86 p<0.05) were independently associated with low VO$_2$AT (<10mls/kg/min). However, jaundice was not associated with low VO$_2$AT. These results are shown in Table \ref{table:cpet_oj_table5}

\input{Tables/c3_t5}

\subsubsection{Scatter-plot analysis}
Scatter-plot analysis comparing serum bilirubin and VO$_2$AT as continuous variables is depicted in Figure 1. This shows that the relationship between serum bilirubin and AT is weak with an r2 value of only 0.04 (I will have to confirm this but it is not more than 0.1).

@@

While the aforementioned studies have undermined the role of PBD in jaundiced patients undergoing pancreaticoduodenectomy, the results of the present study show for the first time that the premise for performing PBD, namely the adverse effect of jaundice on cardiopulmonary physiology may itself be flawed in patients undergoing pancreaticoduodenectomy. In our study, obstructive jaundice including severe obstructive jaundice did not affect cardiopulmonary exercise capacity as measured by VO$_2$AT or the peak oxygen consumption. These findings taken together with previously published findings of adverse effects of PBD further support the fact that major surgery may be safe in jaundiced patients without subjecting them to preoperative biliary drainage. 
The basis of the relationship between low VO$_2$AT and raised BMI is not clear.

However, such an association has been previously reported.\parencite{horwich_relationship_2009}  This may reflect the difficulty in obtaining accurate VO$_2$AT values in obese patients as a result of the calculations involved rather than due to true cardiopulmonary dysfunction. Other authors have suggested that different thresholds for CPET parameters may have to be considered in obese patients to improve risk-prediction.\parencite{donnelly_criteria_1990,hulens_exercise_2001} Cardiopulmonary exercise testing measures oxygen delivery to skeletal muscle. Adipose tissue, however, does not contribute to the metabolic activity that is measured during CPET. However, AT as normally reported, is calculated by dividing the oxygen consumption per minute at the ‘anaerobic threshold' into the weight of the patient. However, this does not account for the disproportionately higher amount of adipose tissue in overweight/obese patients resulting in a spuriously low AT (in mls/kg/min). The present study found no association between cardiorespiratory comorbidity and VO$_2$AT. Low VO$_2$AT in female patients and overweight/obese patients should be interpreted with caution as this may not be due to true poor aerobic capacity.

>>>>

\lhead{Chapter \ref{ch_cpet_jaundice}. \emph{Obstructive Jaundice}} % This is for the header on each page - perhaps a shortened title

@@

\section{Patients and Methods}
Patients who underwent classical or pylorus-preserving pancreaticoduodenectomy for periampullary lesions (both benign and malignant) between August 2008 and April 2013 and had undergone cardiopulmonary exercise testing as part of their preoperative work-up at the West of Scotland Pancreatic Unit, Glasgow Royal Infirmary, Glasgow were included in the study. Established criteria for resectability in patients with malignant disease were used as outlined in previous published work. Segmental or wedge resection of the portal vein or superior mesenteric vein was carried out if the lesion was otherwise resectable.

@@

\subsection{Obstructive Jaundice}
Serum bilirubin levels were measured in all patients on the day before surgery. Obstructive jaundice was defined as bilirubin levels greater than 35 micromol/litre and severe obstructive jaundice was defined as bilirubin levels greater than 250 micromol/litre. This threshold was selected because the DROP trial did not investigate patients with bilirubin levels greater than 250 micromol/litre and this study aimed to evaluate preoperative pathophysiology in this particular group. 
\todo{A word about stents}

\subsection{Cardiopulmonary Exercise Test}
Cardiopulmonary exercise tests were performed in the Department of Respiratory Medicine at the Glasgow Royal Infirmary using the ZAN-600 CPET suite (nSpire Health, Longmont, CO 80501, USA) (9). All patients underwent standard pulmonary function tests and spirometry prior to cardiopulmonary exercise testing. A cycle ergometer was used to perform a symptom-limited, incremental work-load test preceded by a 3-minute rest period. The test was stopped when patients achieved their maximum exercise tolerance, when significant ischaemic changes occurred on ECG or for other safety reasons. Peak oxygen consumption achieved at this stage was defined as $\dot{V}_{O_2}$Peak. The $\dot{V}_{O_2}$AT was calculated using the V-slope\parencite{beaver_new_1986,sue_metabolic_1988} and ventilatory equivalents\parencite{society_ats/accp_2003} methods. $\dot{V}_{O_2}$AT less than 10 ml/kg/min was considered to be low based on previous work by us\parencite{chandrabalan_pre-operative_2013} as well as Ausania and co-workers\parencite{ausania_effects_2012} which has shown increased incidence of complications in these patients. Oxygen consumption at peak exercise ($\dot{V}_{O_2}$Peak) was dichotomised using a cut-off of 16 ml/kg/min. Detailed description of cardiopulmonary exercise testing as well as the physiological parameters described in this study are published elsewhere.\parencite{balady_clinicians_2010}

\subsection{Statistics}
Grouping of the variables was carried out using standard or previously published thresholds. In the absence of such thresholds, the variables were treated as continuous variables. Non-parametric tests were used to analyse the association between categorical and continuous variables while Chi-square tests were used to analyse the association between categorical variables. Univariate and multivariate binary logistic regression analysis was used to study the relationship between preoperative patient characteristics and $\dot{V}_{O_2}$AT / $\dot{V}_{O_2}$Peak. SPSS software (Version 17.0; SPSS Inc., Chicago, IL, USA) was used to perform statistical analysis.

@@

\section{Results}
One-hundred and thirty eight patients had undergone pancreaticoduodenectomy (n=138), with preoperative cardiopulmonary exercise testing during the study period. Over half the patients were male (n=93, 67\%). Approximately half the number of patients were over the age of 65 (n=68, 49\%) and overweight or obese (n=69, 50\%). Cardiovascular comorbidity was present in 58 patients (42\%) and respiratory comorbidity was present in 12 patients (9\%). Fifty patients (36\%) had a history of cigarette smoking. The POSSUM Physiology Score was greater than 14 in 61 patients (44\%). Obstructive jaundice (serum bilirubin 35 – 250) was present in 32 (23\%) patients while severe obstructive jaundice (serum bilirubin > 250) was present in 19 (14\%) patients. 
The baseline demographic and clinical characteristics of non-jaundiced and jaundiced patients are shown in Table \ref{table:cpet_oj_patient}. A larger proportion of jaundiced patients were females compared to the non-jaundiced cohort (p$<$0.05) and smokers (p$<$0.05). Patients with jaundice were more likely to have an elevated POSSUM Physiology Score (p$<$0.005). Patients with cancer were more likely to be jaundiced (p$<$0.001). However, there was no statistically significant difference in age, BMI, cardiovascular comorbidity, or respiratory comorbidity between the non-jaundiced and jaundiced patients.

\input{Tables/cpet_oj_patient}

The relationship between obstructive jaundice and preoperative blood tests is shown in Table \ref{table:cpet_oj_bloods}. While obstructive jaundice was statistically associated with multiple haematological and biochemical abnormalities, most of these did not appear to be of clinically significance. As expected, obstructive jaundice was associated with markedly elevated liver enzymes with severity of derangement associated with severity of jaundice. Obstructive jaundice and sOJ were associated with increasing CRP levels (p$<$0.001) and decreasing serum albumin levels (p$<$0.001). Obstructive jaundice was not associated with deranged renal function with both urea and creatinine remaining similar across all cohorts (p=0.09 and p=0.22 respectively).

\input{Tables/cpet_oj_bloods}

There was no association between obstructive jaundice and preoperative pulmonary function tests (Table \ref{table:cpet_oj_pft}).

\input{Tables/cpet_oj_pft}

\subsubsection{Univariate analysis of obstructive jaundice versus CPET}
The relationship between obstructive jaundice and multiple physiological parameters measured at cardiopulmonary exercise testing is shown in Table \ref{table:cpet_oj_anaerobic}. There was an inverse relationship between oxygen consumption at the anaerobic threshold ($\dot{V}_{O_2}$AT) and increasing severity of jaundice (p$<$0.05). However, no such linear relationship was noted between any of the other parameters measured both at anaerobic threshold and at peak exercise in spite of apparent statistically significant associations.

\input{Tables/cpet_oj_anaerobic}
\input{Tables/cpet_oj_peak}

\subsubsection{Association between preoperative clinico-pathological factors and $\dot{V}_{O_2}$AT}
On multivariate analysis female sex (HR 3.75 CI 1.57-8.95 p$<$0.005), high BMI (HR 3.65 CI 1.61-8.26 p$<$0.005), presence of cancer (HR 4.02 CI 1.33-12.16 p$<$0.05) and raised CRP (HR 2.98 CI 1.29-6.86 p$<$0.05) were independently associated with low $\dot{V}_{O_2}$AT ($<$10mls/kg/min). However, jaundice was not associated with low $\dot{V}_{O_2}$AT. These results are shown in Table \ref{table:cpet_oj_at_regression}

\input{Tables/cpet_oj_at_regression}
\input{Tables/cpet_oj_peak_regression}

\subsubsection{Scatter-plot analysis}
Scatter-plot analysis comparing serum bilirubin and $\dot{V}_{O_2}$AT as continuous variables is depicted in Figure 1. This shows that the relationship between serum bilirubin and AT is weak with an r2 value of only 0.04 (I will have to confirm this but it is not more than 0.1).

@@

While the aforementioned studies have undermined the role of PBD in jaundiced patients undergoing pancreaticoduodenectomy, the results of the present study show for the first time that the premise for performing PBD, namely the adverse effect of jaundice on cardiopulmonary physiology may itself be flawed in patients undergoing pancreaticoduodenectomy. In our study, obstructive jaundice including severe obstructive jaundice did not affect cardiopulmonary exercise capacity as measured by $\dot{V}_{O_2}$AT or the peak oxygen consumption. These findings taken together with previously published findings of adverse effects of PBD further support the fact that major surgery may be safe in jaundiced patients without subjecting them to preoperative biliary drainage. 
The basis of the relationship between low $\dot{V}_{O_2}$AT and raised BMI is not clear.

However, such an association has been previously reported.\parencite{horwich_relationship_2009}  This may reflect the difficulty in obtaining accurate $\dot{V}_{O_2}$AT values in obese patients as a result of the calculations involved rather than due to true cardiopulmonary dysfunction. Other authors have suggested that different thresholds for CPET parameters may have to be considered in obese patients to improve risk-prediction.\parencite{donnelly_criteria_1990,hulens_exercise_2001} Cardiopulmonary exercise testing measures oxygen delivery to skeletal muscle. Adipose tissue, however, does not contribute to the metabolic activity that is measured during CPET. However, AT as normally reported, is calculated by dividing the oxygen consumption per minute at the ‘anaerobic threshold' into the weight of the patient. However, this does not account for the disproportionately higher amount of adipose tissue in overweight/obese patients resulting in a spuriously low AT (in mls/kg/min). The present study found no association between cardiorespiratory comorbidity and $\dot{V}_{O_2}$AT. Low $\dot{V}_{O_2}$AT in female patients and overweight/obese patients should be interpreted with caution as this may not be due to true poor aerobic capacity.
